---
title: "Results"
author: "LucianoRogerio and Jean-Luc Jannink"
date: "2023-07-30"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Setup.  
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = F,
  tidy.opts=list(strict=FALSE,width.cutoff=100), highlight=TRUE)
rm(list=ls())
library(tidyverse); library(magrittr); library(genomicMateSelectR)

netID <- "jj332"
jobName <- "NRCRI_2023GS"
outputFileNames <- readRDS(paste0("~/", jobName, "/output/OutputFileNames.rds"))
for (i in 1:length(outputFileNames))
  assign(names(outputFileNames)[i], outputFileNames[i])
```

## Raw data.  
NOTE: All other .Rmd files copy files over from /home/netID to the local
workdir, set here::here to the local workdir, then source data locally.  
For these results, the assumption is that all previous .Rmd scripts have saved
their results over to /home/netID and that those results should be pulled from
there directly.  

Summary of the number of unique plots, locations, years, etc. in the cleaned plot-basis data. See here for details..
```{r Raw data Info, warning = FALSE, error = FALSE}
rawdata <- readRDS(paste0("/home/", netID, "/" , jobName,
                          "/output/", exptDesigns))
rawdata %>% 
  summarise(Nplots=nrow(.),
            across(c(locationName,studyYear,studyName,TrialType,GID), ~length(unique(.)),.names = "N_{.col}")) %>% 
  rmarkdown::paged_table()
```

8793 unique clone names in the phenotype data, across >57K plots.

This is not the same number of clones as are expected to be genotyped-and-phenotyped.

Break down the plots based on the trial design and TrialType (really a grouping of the population that is breeding program specific), captured by two logical variables, CompleteBlocks and IncompleteBlocks.

```{r Experimental Designs}
rawdata %>% 
  count(TrialType,CompleteBlocks,IncompleteBlocks) %>% 
  spread(TrialType,n) %>% 
  rmarkdown::paged_table()
```

Next, look at breakdown of plots by TrialType (rows) and locations (columns):

```{r Info Trial Type vs Locations}
rawdata %>% 
  count(locationName,TrialType) %>% 
  spread(locationName,n) %>% 
  rmarkdown::paged_table()
```

```{r Traits Distribution}
traits <- c("DMC", "DRY", "PLTHT", "BRNHT1", "logFYLD",
          "CGMS1", "CGMS2", "MCMDS", "DMCsg", "IVig1-7",
          "StC", "HI", "logDRYsgLD", "logTOPYLD")
rawdata %>% 
  select(locationName,studyYear,studyName,TrialType,any_of(traits)) %>% 
  pivot_longer(cols = any_of(traits), values_to = "Value", names_to = "Trait") %>% 
  ggplot(.,aes(x=Value,fill=Trait)) + geom_histogram() + facet_wrap(~Trait, scales='free') + 
  theme_bw() + scale_fill_viridis_d() + 
  labs(title = "Distribution of Raw Phenotypic Values")
```

How many genotyped-and-phenotyped clones?

```{r Clones Genotyped and phenotyped}
rawdata %>% 
  select(any_of(c("locationName","studyYear","studyName","TrialType",
                "germplasmName","FullSampleName","GID")),
         any_of(traits)) %>% 
  pivot_longer(cols = any_of(traits), values_to = "Value", names_to = "Trait") %>%
  filter(!is.na(Value)) %>%
  distinct(germplasmName, GID) %>% 
  rmarkdown::paged_table()
```
There are 2,083 genotyped-and-phenotyped clones!

Table of germplasmName-DNA-sample-name matches are here: output/OnlyChosen_germplasmName_to_FullSampleName_matches_NRCRI_2022Jul09.csv.

List of DNA-sample-names are here:

1. RefPanel (containing NRCRI TP):
2. New samples (DCAs22-7085):

## BLUPs

These are the BLUPs combining data for each clone across trials/locations without genomic information, used as input for genomic prediction downstream.

```{r BLUPs Info}
blups <- readRDS(paste0("/home/", netID, "/" , jobName,
                        "/output/", blupsForTraining)) 

blups %>% 
  left_join(nestDesignsDetectedByTraits(rawdata,traits) %>% 
  mutate(Nplots=map_dbl(MultiTrialTraitData,nrow)) %>% 
    select(Trait,Nplots)) %>% 
  mutate(Nclones=map_dbl(blups,~nrow(.)),
         NoutliersRemoved=map2_dbl(outliers1,outliers2,~length(.x)+length(.y))) %>% 
  #relocate(c(Nclones,NoutliersRemoved),.after = Trait) %>% 
  #select(-blups,-varcomp,-outliers1,-outliers2) %>% 
  select(Trait,Nplots,Nclones,NoutliersRemoved,Vg,Ve,H2) %>% 
  mutate(across(is.numeric,~round(.,4))) %>% arrange(desc(H2)) %>% 
  rmarkdown::paged_table()

blups %>% 
  select(Trait,blups) %>% 
  unnest(blups) %>% 
  ggplot(.,aes(x=drgBLUP,fill=Trait)) + geom_histogram() + facet_wrap(~Trait, scales='free') + 
  theme_bw() + scale_fill_viridis_d() + 
  labs(title = "Distribution of de-regressed BLUP Values")

blups %>% 
  select(Trait,blups) %>% 
  unnest(blups) %>% 
  ggplot(.,aes(x=Trait,y=REL,fill=Trait)) + geom_boxplot(notch=T) + #facet_wrap(~Trait, scales='free') + 
  theme_bw() + scale_fill_viridis_d() + theme(axis.text.x = element_text(angle=90)) +
  labs(title = "Distribution of BLUP Reliabilities")
```

## Marker density and distribution

```{r Marker density}
snps <- readRDS(paste0("/home/", netID, "/" , jobName,
                       "/output/", dosMatPred))
mrks <- colnames(snps) %>% 
  tibble(SNP_ID=.) %>% 
  separate(SNP_ID,c("Chr","Pos","Allele"),"_") %>% 
  mutate(Chr=as.integer(gsub("S","",Chr)),
         Pos=as.numeric(Pos))
mrks %>% 
  ggplot(.,aes(x=Pos,fill=as.character(Chr))) + geom_histogram() + 
  facet_wrap(~Chr,scales = 'free') + theme_bw() + 
  scale_fill_viridis_d() + theme(legend.position = 'none',axis.text.x = element_text(angle=90))
mrks %>% count(Chr) %>% rmarkdown::paged_table()
```

## Prediction accuracy

Check prediction accuracy: Evaluate prediction accuracy with cross-validation.

```{r Prediction Accuracy}
cv <- readRDS(paste0("/home/", netID, "/" , jobName,
                     "/output/", crossValResults)) %>% 
#  bind_rows(readRDS(here::here("output","cvresults_ADE_1_2022Jul10.rds"))) %>% 
#  bind_rows(readRDS(here::here("output","cvresults_ADE_2_2022Jul10.rds"))) %>% 
  unnest(accuracyEstOut) %>% 
  dplyr::select(-splits)

# cv %<>% 
#  mutate(Trait=factor(Trait,levels=traits),
#         modelType=factor(modelType,levels=c("A","ADE")))
```


## Table of mean accuracies


5-fold cross-validation, replicated 20 times.

Mean accuracy and upper/lower 95% interval.

Two prediction models: A (additive-only) and ADE (additive + dominance + additive-by-dominance epistasis).

```{r Mean cross validation accuracies}
cv %>% 
  group_by(Trait,modelType) %>% 
  # use accGETGV. For modelA we GETGV==GEBV. For modelADE we don't want GEBV, just GETGV.
  # 30 July 2023: don't have accGETGV because didn't do it yet...
  summarize(meanAccuracy=mean(Accuracy,na.rm=T),
            lower5pct=quantile(Accuracy,probs = c(0.05),na.rm=T),
            upper5pct=quantile(Accuracy,probs = c(0.95),na.rm=T)) %>% 
  mutate(across(is.numeric,~round(.,2))) %>% arrange(modelType,desc(meanAccuracy)) %>% 
  rmarkdown::paged_table()
```

## Boxplot of accuracies


5-fold cross-validation, replicated 20 times.

Two prediction models: A (additive-only) and ADE (additive + dominance + additive-by-dominance epistasis).


```{r Accuracy boxplots}
# Again, no accGETGV, only Accuracy
cv %>% 
    ggplot(.,aes(x=Trait,y=Accuracy,fill=modelType)) + 
  geom_boxplot(position = "dodge2",color='gray50',size=0.5, notch = T) + 
  geom_hline(yintercept = 0, color='darkred') + 
  theme_bw() + 
  theme(strip.text.x = element_text(face='bold', size=12),
        axis.text.y = element_text(face='bold', size=14, angle = 0),
        axis.text.x = element_text(face='bold', size=18, angle = 90, hjust = 1),
        axis.title.y = element_text(face='bold', size=12),
        plot.title = element_text(face='bold'),
        legend.text = element_text(face='bold',size=16),
        legend.title = element_text(face='bold',size=16),
        legend.position = 'bottom') + 
  scale_fill_viridis_d() + 
  labs(title="Prediction Accuracies", y="GEBV or GETGV accuracy",x=NULL) +
  geom_hline(yintercept = 0, color='darkred')
```

1. Accuracy estimates are most improved relative to previously. I didn’t run the precise cross-validation folds so the judgement is based on visual comparison to the Dec. 2020 plot.
2. DYLD and FYLD are not well predicted and I would not recommend using them based on selection.


## Genetic Gain

```{r Genetic Gain}
gebvs <- read.csv(
  paste0("~/", jobName, "/output/GEBV_NRCRI_ModelA_2023Jul30.csv"),
  stringsAsFactors = F) %>% 
  pivot_longer(cols = any_of(traits),names_to="Trait", values_to="GEBV")

gebvs %<>% 
  mutate(Trait=factor(Trait,levels=traits),
         Group=factor(Group,levels=c("nrTP","C1a","C1b","C2a","C2b","C3a","C3b", "C4a","C4b")))

gebvs %>% 
  group_by(Trait, Group) %>% 
  summarize(meanGEBV = mean(GEBV), 
            stdErr = sd(GEBV)/sqrt(n()), 
            upperSE = meanGEBV + stdErr, 
            lowerSE = meanGEBV - stdErr) %>% 
  ggplot(., aes(x = Group, 
                y = meanGEBV, 
                fill = Trait)) + 
  geom_bar(stat = "identity", color = "gray60", 
           size = 1.25) + 
  geom_linerange(aes(ymax=upperSE, ymin=lowerSE), color="gray60", size = 1.25) + 
  facet_wrap(~Trait, scales = "free") + 
  theme_bw() + 
  geom_hline(yintercept = 0, size = 1.15, color = "black") + 
  theme(axis.text.x = element_text(face = "bold", angle = 90, size = 12), 
        axis.title.y = element_text(face = "bold", size = 14), 
        legend.position = "none", 
        strip.background.x = element_blank(), 
        strip.text = element_text(face = "bold", size = 14)) + 
  scale_fill_viridis_d() + 
  labs(x = NULL, y = "Mean GEBVs")
```

